---
title: "Reproducible Research: Peer Assessment 1"
author: "Sherilyn"
date: "17 May 2017"
output:  md_document
   
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


Loading and preprocessing the data

```{r loaddata}
library(ggplot2)
library(dplyr)
library(data.table)
library (lubridate)
library(knitr)
setwd("c:/courseraprojects")
activity <- read.csv("activity.csv")

```


Make a histogram on total number of steps per day

```{r}
aggsteps<-aggregate(steps~date, data=activity, FUN = sum)
hist(aggsteps$steps,xlab = "Number of steps", main = "Total Number of steps per day")
```

Calculate and report the mean and median total number of steps taken per day

mean
```{r}
mean(aggsteps$steps)
```

median
```{r}
median(aggsteps$steps)
```


What is the average daily activity pattern? Make a time series plot (i.e.type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
averages <- aggregate(x = list(steps = activity$steps), by = list(interval = activity$interval), 
                      FUN = mean, na.rm = TRUE)
ggplot(data = averages, aes(x = interval, y = steps)) + geom_line() + xlab("5-minute interval") + 
        ylab("average number of steps taken")
```


Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
averages[which.max(averages$steps), ]
```


Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs).

```{r}
na<-is.na(activity$steps)
table(na)
```


Inputing missing values,NA, for days/intervals

```{r}
fill.value <- function(steps, interval) {
        filled <- NA
        if (!is.na(steps)) 
                filled <- c(steps) else filled <- (averages[averages$interval == interval, "steps"])
                return(filled)
}
filled.data <- activity
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
```


Make a histogram of the total number of steps taken each day and calculate the mean and median total number of steps.

```{r}
total.steps <- tapply(filled.data$steps, filled.data$date, FUN = sum)
qplot(total.steps, binwidth = 1000, xlab = "total number of steps taken each day")

mean(total.steps)
median(total.steps)
```


Are there differences in activity patterns between weekdays and weekends? 

```{r}
dayweek<-function(date){
        day<-weekdays(date)
        if(day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"))
                return("weekday") else if (day %in% c("Saturday","Sunday"))
                        return("weekend") else stop ("invalid date")}
        
filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN = dayweek) 
```

        
Make a panel plot containing plots of average number of steps taken on weekdays and weekends. 

```{r}
averages<-aggregate(steps~interval + day, data = filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) + 
        xlab("5-minute interval") + ylab("Number of steps")
```


